Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-15 paper/tv.tex
m

\lstset{
   language=JavaScript,
   %%backgroundcolor=\color{lightgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
%frame=single
numbers=left,   
xleftmargin=1.5em, 
    numberstyle=\tiny\color{mGray},
   numbersep=10pt,
   tabsize=4,
   breaklines=true,
   showtabs=false,
   captionpos=b,
   keywordstyle=[2]\color{purple}\bfseries, %
   morekeywords=[2]{do, hiphop,every, loop,suspend, when, signal, end, present, nothing, emit,in, out, yield,abort, fork, par, run},
}

@@

%(\code{\{Go\} \Rightarrow \_})
%(\code{\{Ready\} \Rightarrow \_}, \code{t_L^2 \Rightarrow true})
Termination is guaranteed because the set of derivatives to be considered is finite, and possible cycles are detected using 
\emph{memorization} \cite{DBLP:conf/tableaux/Brotherston05}, cf., \tabref{tab:reoccur}.

\subsection{The Fischer's mutual exclusion algorithm}

\begin{wrapfigure}{L}{0.53\columnwidth}
  \vspace{-1mm}
\begin{lstlisting}
var x := -1; 
var ct := 0;
void proc (int i) {
  [x=-1] 
  deadline(event["Update"(i)]{x:=i},d1);
  delay (d2);
  if (x=i) {
    event["Critical"(i)]{ct:=ct+1};
    event["Exit"(i)]{ct:=ct-1; x:=-1};
    proc (i);
  } else {
    proc (i);}}
void main () { 
  proc(0) || proc(1) || proc(2); }
\end{lstlisting}  
  \vspace{-1mm}
  \caption{Fischer's mutual exclusion algorithm.}\label{fig:overview_ficher}
     \vspace{-1mm}
\end{wrapfigure}

The following models Fischer's mutual exclusion algorithm. 
where d1 and d2 are two integer constants with d1 < d2; 
x and ct are global variables. The protocol is modeled as 
process Protocol, which is the parallel composition of three 
processes. Each of the three processes attempts to enter 
the critical section when x is -1, i.e. no other process is 
currently attempting. Once the process is active, it sets x 
to its identity i within d1 time units (captured by deadline[d1]). 
Then it idles for d2 time units (captured by Wait[d2]) and then 
checks whether x is still i. If so, it enters the critical 
section and leaves later. Otherwise, it restarts from the 
beginning.

We show that our method is able to prove the algorithm in the 
section of implementation.  prove the algorithm in the 
section of implementation.

>>>>

\lstset{
   language=JavaScript,
   %%backgroundcolor=\color{lightgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
%frame=single
numbers=left,   
xleftmargin=1em, 
    numberstyle=\tiny\color{mGray},
   numbersep=10pt,
   tabsize=4,
   breaklines=true,
   showtabs=false,
   captionpos=b,
   keywordstyle=[2]\color{purple}\bfseries, %
   morekeywords=[2]{do, hiphop,every, loop,suspend, when, signal, end, present, nothing, emit,in, out, yield,abort, fork, par, run},
}

@@

%(\code{\{Go\} \Rightarrow \_})
%(\code{\{Ready\} \Rightarrow \_}, \code{t_L^2 \Rightarrow true})
Termination is guaranteed because the set of derivatives to be considered is finite, and possible cycles are detected using 
\emph{memorization} \cite{DBLP:conf/tableaux/Brotherston05}, cf., \tabref{tab:reoccur}. 
\\

\subsection{Verifying the Fischer’s Mutual Exclusion Protocol}

\begin{wrapfigure}{L}{0.51\columnwidth}
  \vspace{-5mm}
\begin{lstlisting}
var x := -1; 
var ct:= -1;

void proc (int i) {
  [x=-1] 
  deadline(event["Update"(i)]{x:=i},d1);
  delay (d2);
  if (x=i) {
    event["Critical"(i)]{ct:=i};
    event["Exit"(i)]{ct:=-1; x:=-1};
    proc (i);
  } else {
    proc (i);}}

void main () 
/*req: d1<d2/\emp
  ens: true /\ (([ct=-1]\/[ct=x])^*) */ 
{ proc(0) || proc(1) || proc(2); }
\end{lstlisting}  
  %\vspace{-1mm}
  \caption{Fischer's mutual exclusion algorithm.}\label{fig:overview_ficher}
    % \vspace{-2mm}
\end{wrapfigure}

\figref{fig:overview_ficher} presents the classical 
Fischer's mutual exclusion algorithm. 
where \codem{d1} and \codem{d2} are two integer constants
 with \codem{d1<d2}; 
\code{x} and \code{ct} are global variables. The protocol is 
modeled as 
process protocol, which is the parallel composition of three 
processes. Each of the three processes attempts to enter 
the critical section when \code{x} is \code{-1}, i.e. no 
other process is 
currently attempting. Once the process is active, it sets \codeme{}{x} 
to its identity \codem{i} within \codem{d1} time units (captured by \codem{deadline(..., d1}). 
Then it idles for \codem{d2} time units (captured by \codem{delay(d2)}) and then 
checks whether \codem{x} is still \codem{i}. If so, 
it enters the critical 
section and leaves later. Otherwise, it restarts from the 
beginning.

We show the specification of \codem{main} using \timedEffects, which
restricts the relation between \codem{d1} and \codem{d2}. When 
\codem{d1<d2} is provided, it ensures a repeated trace where the value of 
\codem{ct} is either \codem{-1} or equals to \codem{x}. The specification
implies that this implementation is indeed mutual exclusive. 
Our prototype system is able to soundly check such time-critical 
algorithms. \autoref{sec:Evaluation} presents the details about the implementation and evaluation.
